---
title: "Assignment 2 - Language Development in ASD - Making predictions"
author: "Gustav, Magnus, Morten, Nicoline, Tobias"
date: "August 9, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Exercise 1) Testing model performance

How did your models from last time perform? In this exercise you have to compare the results on the training data and on the test data. Report both of them. Compare them. Discuss why they are different.

- recreate the models you chose last time (just write the model code again and apply it to your training data (from the first assignment))
- calculate performance of the model on the training data: root mean square error is a good measure. (Tip: google the function rmse())
- create the test dataset (apply the code from assignment 1 to clean up the 3 test datasets)
- test the performance of the models on the test data (Tips: google the functions "predict()")
- optional: predictions are never certain, can you identify the uncertainty of the predictions? (e.g. google predictinterval())


```{r, include = FALSE}

pacman::p_load(readr,dplyr,stringr,lmerTest,Metrics,caret, tidyverse, merTools)

## Clean up function, included to inspire you

CleanUpData <- function(Demo,LU,Word){
  
  Speech <- merge(LU, Word) %>% 
    rename(
      Child.ID = SUBJ, 
      Visit = VISIT) %>%
    mutate(
      Visit = as.numeric(str_extract(Visit, "\\d")),
      Child.ID = gsub("\\.","", Child.ID)
      ) %>%
    dplyr::select(
      Child.ID, Visit, MOT_MLU, CHI_MLU, types_MOT, types_CHI, tokens_MOT, tokens_CHI
    )
  
  Demo <- Demo %>%
    dplyr::select(
      Child.ID, Visit, Ethnicity, Diagnosis, Gender, Age, ADOS, MullenRaw, ExpressiveLangRaw, Socialization
    ) %>%
    mutate(
      Child.ID = gsub("\\.","", Child.ID)
    )
    
  Data = merge(Demo, Speech, all = T)
  
  Data1 = Data %>% 
     subset(Visit == "1") %>% 
     dplyr::select(Child.ID, ADOS, ExpressiveLangRaw, MullenRaw, Socialization) %>%
     rename(Ados1 = ADOS, 
            verbalIQ1 = ExpressiveLangRaw, 
            nonVerbalIQ1 = MullenRaw,
            Socialization1 = Socialization) 
  
  Data = merge(Data, Data1, all = T) %>%
    mutate(
      Child.ID = as.numeric(as.factor(as.character(Child.ID))),
      Visit = as.numeric(as.character(Visit)),
      Gender = recode(Gender, 
         "1" = "M",
         "2" = "F"),
      Diagnosis = recode(Diagnosis,
         "A"  = "ASD",
         "B"  = "TD")
    )

  return(Data)
}

# Load training Data (Magnus)
df <- read_csv('autisme')
df$X1 <- NULL
df <- df %>% filter(!is.na(CHI_MLU))
#- recreate the models you chose last time (just write the code again and apply it to Train Data) (Gustav)
m <- lmerTest::lmer(CHI_MLU ~ VISIT * Diagnosis + (1 | SUBJ), df, REML = F)
m <- update(m, .~. + I(VISIT^2))

#- calculate performance of the model on the training data: root mean square error is a good measure. (Tip: google the function rmse()) (Morten)


pre <- predict(m, newdata = df, allow.new.levels = T)

rmse(df$CHI_MLU, pre)
#- create the test dataset (apply the code from assignment 1 or my function to clean up the 3 test datasets)
# Test data (Nicoline)
LU <- read_csv('LU_test.csv')
Demo <- read_csv('demo_test.csv')
Word <- read_csv('token_test.csv')

clean <- CleanUpData(Demo, LU, Word)

colnames(clean)[1] <- 'SUBJ'
colnames(clean)[2] <- 'VISIT'

clean <- clean %>% filter(!is.na(clean$CHI_MLU))

#- test the performance of the models on the test data (Tips: google the functions "predict()") (Tobias)
pre2 <- predict(m, newdata = clean)

rmse(clean$CHI_MLU, pre2)
#- optional: predictions are never certain, can you identify the uncertainty of the predictions? (e.g. google predictinterval()) (Magnus)
prei <- predictInterval(m, newdata = clean)

summary(prei)

```

### Exercise 2) Model Selection via Cross-validation (N.B: ChildMLU!)

One way to reduce bad surprises when testing a model on new data is to train the model via cross-validation. 

In this exercise you have to use cross-validation to calculate the predictive error of your models and use this predictive error to select the best possible model.

- Use cross-validation to compare your model from last week with the basic model (Child MLU as a function of Time and Diagnosis, and don't forget the random effects!)
- (Tips): google the function "createFolds";  loop through each fold, train both models on the other folds and test them on the fold)


- Now try to find the best possible predictive model of ChildMLU, that is, the one that produces the best cross-validated results.

- Bonus Question 1: What is the effect of changing the number of folds? Can you plot RMSE as a function of number of folds?
- Bonus Question 2: compare the cross-validated predictive error against the actual predictive error on the test data


```{r}
#- Create the basic model of ChildMLU as a function of Time and Diagnosis (don't forget the random effects!). (Gustav)

m1 <- CHI_MLU ~ VISIT + Diagnosis + (1|SUBJ)

#- Make a cross-validated version of the model. (Tips: google the function "createFolds";  loop through each fold, train a model on the other folds and test it on the fold) (Morten)

#function for cross-validating models taking inputs x = model, y = number of folds, z = name of model
cross_validate <- function(x, y, z) {
  k <- y
  folds <- createFolds(unique(df$SUBJ), k = k, list = T, returnTrain = F)
  trainrmse <- rep(NA, k)
  testrmse <- rep(NA, k)
  i = 1
  for (fold in folds) {
    train = subset(df, !(SUBJ %in% fold))
    test = subset(df, SUBJ %in% fold)
    model = lmer(x, train, REML = FALSE)
    test$prediction = predict(model, test, allow.new.levels = T)
    train$prediction = fitted(model)
    trainrmse[i] = rmse(train$CHI_MLU, fitted(model))
    testrmse[i] = rmse(test$CHI_MLU, test$prediction)
    i = i + 1
  }
  rm <- data.frame("model" = z, "train" = trainrmse, "test" = testrmse, "diff" = testrmse - trainrmse, "meantrain" = rep(mean(trainrmse), k), "meantest" = rep(mean(testrmse), k), "meandiff" = rep(mean(testrmse) - mean(trainrmse), k), "sdtrain" = rep(sd(trainrmse), k), "sdtest" = rep(sd(testrmse), k))
  return(rm)
}

m1_rmse <- cross_validate(m1, 6, "m1")

#- Now try to find the best possible predictive model of ChildMLU, that is, the one that produces the best cross-validated results. (Nicoline)

m2 <- CHI_MLU ~ VISIT * Diagnosis + (1|SUBJ)

m2_rmse <- cross_validate(m2, 6, "m2")

m3 <- CHI_MLU ~ VISIT * Diagnosis + I(VISIT^2) + (1|SUBJ)

m3_rmse <- cross_validate(m3, 6, "m3")

m4 <- CHI_MLU ~ VISIT + Diagnosis + types_CHI + MOT_MLU + (1|SUBJ)

m4_rmse <- cross_validate(m4, 6, "m4")

m5 <- CHI_MLU ~ VISIT + Diagnosis * types_CHI + MOT_MLU + (1|SUBJ)

m5_rmse <- cross_validate(m5, 6, "m5")

m6 <- CHI_MLU ~ VISIT * Diagnosis + types_CHI + MOT_MLU + (1|SUBJ)

m6_rmse <- cross_validate(m6, 6, "m6")

#make data frame containing rmse for test, train, difference between the two, mean rmse of test and train respectively, difference between means and standard deviation of train and test (Tobias)

rmse_values <- rbind(m1_rmse, m2_rmse, m3_rmse, m4_rmse, m5_rmse, m6_rmse)
print(rmse_values)
```

### Exercise 3) Assessing the single child

Let's get to business. This new kiddo - Bernie - has entered your clinic. This child has to be assessed according to his group's average and his expected development.

Bernie is one of the six kids in the test dataset, so make sure to extract that child alone for the following analysis.

You want to evaluate:

- how does the child fare in ChildMLU compared to the average TD child at each visit? Define the distance in terms of absolute difference between this Child and the average TD.

- how does the child fare compared to the model predictions at Visit 6? Is the child below or above expectations? (tip: use the predict() function on Bernie's data only and compare the prediction with the actual performance of the child)

```{r}
#make subset containing only Bernie's data (Gustav)

bernie <- subset(clean,  SUBJ == '2')

#get means of child MLU of Bernie and TD children respectively (Magnus)

TDvsBernie <- df %>% 
  filter(Diagnosis == "TD") %>% 
  group_by(VISIT) %>% 
  summarise(mean = mean(CHI_MLU, na.rm = T)) %>% 
  mutate(bernie$CHI_MLU) %>% 
  mutate(bernie$CHI_MLU - mean)
TDvsBernie

#make best model (Morten)

m7 <- lmer(CHI_MLU ~ VISIT * Diagnosis + types_CHI + MOT_MLU + (1|SUBJ), df, REML = FALSE)

#predict Bernie's data based on model (Nicoline)

pre2 <- predict(m7, bernie)

#compare prediction to actual values (Tobias)

pre2

bernie$CHI_MLU

ASD_TD <- df %>% group_by(VISIT, Diagnosis) %>% summarise(mean(CHI_MLU))
print(ASD_TD)
```